Comparative study of Dask and Spark for neuroimaging pipelines\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Natural Sciences and Engineering Research Council}

### 2019-06-17 latex/paper.tex
Modified histogram application

\subsection{Engines}
Spark v2.4.0 and Dask v1.1.4 was used for the experiments.

\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose big data
framework. It provides in-memory computing, data locality and lazy evaluation. Spark
abstract resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant. Its
high-level API provides optimizations on the transformations made to some data. Since
Spark is written in Scala, this optimization is lessen because it requires a
serialization from Python to Java. The API offered by Spark is flexible and suit most
problem however it differ from Python ecosystem. The standalone scheduler is used for
our experiment. It has a LIFO policy.

\subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
offered by dask. Like Spark, it provides in-memory computing, data locality and lazy
evaluation however it does not have fault resilience. Fundamently it is parallized
lists. Dask Bag parallelizes computations of large collection of generic python
objects~\cite{dask-bag}. Also, it offers optimized algorithm that mimic most of the
toolz library. Our experiment use the Dask Distributed~\cite{dask-distributed}
scheduler. It has a LIFO policy. In theory it should speed-up the pipeline excution
by finishing task computations before starting new ones.

\subsubsection{Dask Delayed~\cite{dask-delayed}} Dask Delayed is a low-level API. It
does not offer the optimized algorithm as the high-level APIs however it allows the
user to parallelize custom function that do not fit in the array, dataframe or
MapReduce framework. This is perfect for more complex function which could easily be
parallelized. Like Dask Bag, it offers in-memory computing, data locality and lack
fault resilience. The Dask Distributed scheduler.

\subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
from Dask. Like the other Dask APIs it benifit of in-memory computing and data
locatlity. Unlike the other APIs, Dask Futures triggers computation immediately. When
the user tries to gather data, if available it will return the result otherwise it
will block until the cimputation is done.

\subsection{Dataset}
For the incrementation and histogram experiment the BigBrain~\cite{Amunts:13} is
used. It is a three-dimensional image of a brain scanned at \SI{40}{\micro\metre}.
The data set store the intensity of each voxel and their position. This results in a
75GB image. It is stored in 125 chunks in the minc~\cite{minc} format. To perform our
experiment we convert the chunks into the nifti format. Also, using sam~\cite{sam} we
combine the chunks and split them into 30, 125 and 750 chunks of equal size; 2.5GB,
0.6GB, and 0.1GB, respectively.

\subsection{Applications}

%% Incrementation
\subsubsection{\textbf{Incrementation}}
As our first experiment to benchmark the frameworks we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
image, increments the intensity value of each voxel by 1 for a specified amount of
iterations and write the result to an NFS as a Nifti image. Incrementing the data
reduce the caching effect. This experiment allows us to study the behavior of the
frameworks when all input are processed independently; i.e.\ each task in the graph
only depends on the previous one (see Figure~\ref{Figure fig:tg-inc}). This mimic the
behavior of analyzing multiple independent subjects in parallel.

@@

\begin{algorithm}[!t]
    \caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}
    \Require{\(x\), a sleep delay in float}
    \Require{\(file\), a file containing a chunk}
    \Require{\(fs\), NFS to save image to.}
    \State{read \(chunk\) from \(file\)}
    \ForEach{\(i \in iterations\)}
        \ForEach{\(chunk \in image\)}
            \State{\(chunk\gets chunk+1\)}
            \State{sleep \(x\)}
        \EndFor
    \EndFor
    \State{save \(chunk\) to \(fs\)}
\end{algorithmic}
\end{algorithm}

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation}\label{fig:tg-inc}
\end{figure}

\subsubsection{Histogram}

@@

%% Material
We compare Dask (\textbf{v1.1.4}) and Spark (\textbf{v2.4.0}) based on two metrics
their scheduling and their makespan performance. All scripts used to perform our
experiments are available at
\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
For our experiments, we used Compute Canada Cloud where each instance has 8 cores,
30GB RAM, 10GB bandwidth and is dedicated.

% Histogram
\subsection{Histogram}
For our second experiment, we find the frequency of the intensity of the voxel in
BigBrain and save them as a tuple in a file. In this experiment, the results depend
on all of the input results (see~\ref{Figure fig:tg-histo}). This allows us to study
the behavior of the framework when tasks are dependent. Also, this application
requires data shuffling thus inter worker communication. We decided to omit the Dask
futures API as it is less realistic for this use case.

% BIDS Example
\subsection{BIDS Example}
Our last experiment the BIDS Example algorithm to compare the frameworks.
% Specify more details and the data set used
% SECTION TO COMPLETE -- Experience not done yet

>>>>

\subsection{Engines}
For all our experiments, we used Spark v2.4.0 and Dask v1.1.4.

\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose Big Data
framework. It provides in-memory computing, data locality and lazy evaluation. Spark
abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant. Its
high-level API provides optimizations for transformations applied data. Since
Spark is written in Scala, this optimization is lessen because it requires a
serialization from Python to Java. The API offered by Spark is flexible and suits most
problems however it differs from Python ecosystem. The standalone scheduler is used for
our experiment. It has a LIFO policy.

\subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
offered by Dask. Like Spark, it provides in-memory computing, data locality and lazy
evaluation however it does not have fault resilience. Fundamently, it is parallized
lists. Dask Bag parallelizes computations of large collection of generic python
objects~\cite{dask-bag}. Also, it offers optimized algorithms that mimic most of the
toolz~\cite{toolz} library. Our experiment use the Dask Distributed~\cite{dask-distributed}
scheduler. It has a LIFO policy. In theory it should speed-up the pipeline excution
by finishing a branch of task computations before starting new ones.

\subsubsection{Dask Delayed~\cite{dask-delayed}} Dask Delayed is a low-level API. It
does not offer the optimized algorithm as the high-level APIs however it allows the
user to parallelize custom function that do not fit in the array, dataframe or
MapReduce framework. This is perfect for more complex function which could easily be
parallelized. Like Dask Bag, it offers in-memory computing, data locality and lack
fault resilience. The Dask Distributed scheduler is used.

\subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
from Dask. Like the other Dask APIs, it benifits of in-memory computing and data
locatlity. However, unlike the other APIs, Dask Futures triggers computation
immediately. When the user tries to gather data if available it will return the
result otherwise it will block until the computation is done. Again, the Dask
Distributed scheduler is used.

\subsection{Dataset}
For the incrementation and histogram experiment the BigBrain~\cite{Amunts:13} is
used. It is a three-dimensional image of a brain scanned at \SI{40}{\micro\metre}.
The data set store the intensity of each voxel and their position. This results in a
75GB image. It is stored in 125 chunks in the \textit{MINC}~\cite{minc} format. To
perform our experiment we convert the chunks into the \textit{NifTI}~\cite{nifti}
format. Also, using \textit{sam}~\cite{sam} we combine the chunks and split them into
30, 125 and 750 chunks of equal size; each of 2.5GB, 0.6GB, and 0.1GB, respectively.

\subsection{Applications}
%% Incrementation
\subsubsection{\textbf{Incrementation}}
As our first experiment to benchmark the frameworks, we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
image, increments the intensity value of each voxel by 1 for a specified amount of
iterations, and write the result to an NFS as a NifTI image. Incrementing the data
reduce the caching effect. This experiment allows us to study the behavior of the
frameworks when all input are processed independently; i.e.\ each task in the graph
only depends on the previous one (see Figure~\ref{Figure fig:tg-inc}). This mimic the
behavior of analyzing multiple independent subjects in parallel.

@@

\begin{algorithm}[!b]
    \caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}
    \Require{\(x\), a sleep delay in float}
    \Require{\(file\), a file containing a chunk}
    \Require{\(fs\), NFS to write image to.}
    \State{Read \(chunk\) from \(file\)}
    \ForEach{\(i \in iterations\)}
        \ForEach{\(chunk \in image\)}
            \State{\(chunk\gets chunk+1\)}
            \State{Sleep \(x\)}
        \EndFor
    \EndFor
    \State{Write \(chunk\) to \(fs\)}
\end{algorithmic}
\end{algorithm}

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation}\label{fig:tg-inc}
\end{figure}

\subsubsection{\textbf{Histogram}}
Our second experiment consist to calculate the histogram of the intensity of the
voxel in BigBrain (see Algorithm~\ref{alg:histogram}). It reads the images from the
nfs, calculate inividual image frequency, aggregate the images frequency together,
and write the result to a file. In this experiment, the result depends on all the
input (see Fig.~\ref{fig:tg-histo}). This allows to study the behavior of the
frameworks when tasks are dependent. Also, this application requires data shuffling
thus inter-worker communication. We decided to omit Dask Futures as it does not had
any value over Dask Delayed in this application.

\begin{algorithm}[!t]
    \caption{Histogram}\label{alg:histogram}
    \begin{algorithmic}
    \Require{\(files\), files containing BigBrain}
    \Require{\(fs\), NFS to save image to.}
    \ForEach{\(file \in files\)}
        \State{Read \(chunk\) from \(file\)}
        \State{{Calculate \(frequency\) of \(chunks\)}}
    \EndFor

    \State{\(histogram\gets\)Aggregate \(frequency\) of each \(file\)}

    \State{Write \(histogram\) to \(fs\)}
    \end{algorithmic}
\end{algorithm}

@@

%% Material
We compare Dask (\textbf{v1.1.4}) and Spark (\textbf{v2.4.0}) based on two metrics
their scheduling and their makespan performance. All scripts used to perform our
experiments are available at
\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
For our experiments, we used Compute Canada Cloud where each instance has 8 vCPUs,
30GB RAM, 10GB bandwidth and is dedicated.
